Experience Report: Haskell in Computational Biology

### 2012-03-09 mrfy_experience_report.tex

 simply implements the Viterbi algorithm for computing the optimal parse 
of a protein sequence onto a hidden Markov model.
Thus, the graphical model of 
a HMMER hidden Markov model is exactly that shown in Figure~\ref{plan7}.
In 
contrast, MRFy builds on the approach of SMURF~\cite{Menke:2010ti} and 
SMURFLite~\cite{Daniels:2012}, our previous work.
Like SMURF and SMURFLite, MRFy 
pays attention to longer-range structural interactions than a simple hidden 
Markov model can understand.
In particular, these approaches ``nail down'' 
certain subsequences of the model, corresponding to the non-local hydrogen-bond 
interactions between beta strands.
Beta strands are regions of protein 
structure that ``stick'' to one another.
Because of their pairwise 
interactions, they are highly conserved over the course of evolution.
Moreover, 
again because of these pairwise interactions, a change in one amino acid in a 
beta strand may require a change in a corresponding, hydrogen-bonded residue in 
another beta strand if the protein is to remain stable.
The architecture of the 
resulting graphical model, derived from the HMMER ``Plan7'' architecture, is 
illustrated in Figure~\ref{mrf}.

>>>>

\emph{[GIT CONFLICT HERE:
 simply implements the Viterbi algorithm for computing the optimal parse 
of a protein sequence onto a hidden Markov model.
Thus, the graphical model of 
a HMMER hidden Markov model is exactly that shown in Figure~\ref{plan7}.
In 
contrast, MRFy builds on the approach of SMURF~\cite{Menke:2010ti} and 
SMURFLite~\cite{Daniels:2012}, our previous work.
Like SMURF and SMURFLite, MRFy 
pays attention to longer-range structural interactions than a simple hidden 
Markov model can understand.
]}

In particular, these approaches ``nail down'' 
certain subsequences of the model, corresponding to the non-local hydrogen-bond 
interactions between beta strands.
Because of their pairwise 
interactions, they are highly conserved over the course of evolution.
Moreover, 
again because of these pairwise interactions, a change in one amino acid in a 
beta strand may require a change in a corresponding, hydrogen-bonded residue in 
another beta strand if the protein is to remain stable.
The architecture of the 
resulting graphical model, derived from the HMMER ``Plan7'' architecture, is 
illustrated in Figure~\ref{mrf}.
